### 2025-09-23 paper/sec/authoring-workflow.tex
Purge pointless paragraph.
Some reorg still needed.

The author initially imports the text and accompanying data into the system, which creates a Fluid
representation of the target document. Initially that representation is just a string literal of the form
\kw{"""..."""}, where the triple quotes are the Fluid syntax for an \emph{interpolated string}, i.e.~a string
literal where expressions of the form \kw{\{$e$\}} are permitted within the string. The \SuggestionAgent
analyses the inputs and identifies candidate fragments of natural language which could be replaced by Fluid
expressions. The system then enters the human-in-the loop workflow shown in
Figure~\ref{fig:interpretation-agent}, where the author interacts with the \InterpretationAgent. The system is
initially waiting for the author to select a fragment of text $s$ to interpret, a fragment which may or may
not have been highlighted previously by the \SuggestionAgent. The system then generates a candidate Fluid
expression $e$ by prompting the LLM (\secref{prompt-design} below). The expression is validated through a
closed loop program synthesis step which checks that it is well-formed and evaluates to the target fragment
$s$. If this process succeeds, the system replaces the selected substring $s$ with the interpolation
expression \kw{\{$e$\}}, creating a new document configuration. This is the primary successful path through
the workflow and returns the system to the entry state where it is waiting for another top-level interaction
from the author.

Otherwise, the Fluid command-line either returned an error (either a syntax error or runtime error), or
succeeded but $e$ evaluated to some other string $s' \neq s$. Both cases trigger prompt augmentation with an
appropriate error message, and the system retries generation. Failure of the code synthesis step means that,
even after multiple retries with an augmented prompt, $e$ remained invalid or evaluated to $s' \neq s$. If $e$
was invalid (resulting in an error), no remedial action is open to the author. This is considered an
unsuccessful path through the workflow and returns the system to the start state. In the $s' \neq s$ case, the
can choose to manually abort and return to the entry state, or optionally to \emph{revise the goal}, replacing
$s$ with $s'$ in the target document and retaining $e$ as the candidate expression. This is intended to cover
the situation where the author has made a claim which is \emph{incorrect}, but the data set and surrounding
natural language has caused the LLM to synthesise an expression which generates a different value from the one
specified by the user. Once a valid expression is approved, the updated text is spliced back into the
document, and any downstream interactions (e.g. reloading the web page) are checked. This closed-loop design
combines automated synthesis with validation and author oversight, ensuring both correctness and usability of
the generated code.

The user highlights a fragment of interest, which is then sent to the \InterpretationAgent, which generates a
corresponding candidate expression. Internally the \InterpretationAgent uses a compiler-in-the-loop validation
process to ensure the expression is well-formed. The user is then able to accept the expression (in which case
it is incorporated into the program) or reject and leave the text uninterpreted.

@@

\subsection{Paragraph with generated expression}
\label{subsec:paragraph-with-generated-expression}
Figure \ref{fig:fluid-example-paragraph} shows the Fluid code of a Paragraph with the expression generated by
the authoring assistant.

>>>>

\paragraph{Initial configuration.} The author initially imports the text and accompanying data into the
system, which creates a Fluid representation of the target document. Initially that representation is just a
string literal of the form \kw{"""..."""}, where the triple quotes are the Fluid syntax for an
\emph{interpolated string}, i.e.~a string literal where expressions of the form \kw{\{$e$\}} are permitted
within the string. The \SuggestionAgent analyses the inputs and identifies candidate fragments of natural
language which could be replaced by Fluid expressions.

\paragraph{Authoring workflow.} The system then enters the human-in-the loop authoring workflow shown in
Figure~\ref{fig:interpretation-agent}, where the author interacts with the \InterpretationAgent. The system
initially waits for the author to select a fragment of text $s$ to interpret, a fragment which may or may not
have been highlighted previously by the \SuggestionAgent. The system then generates a candidate Fluid
expression $e$ by prompting the LLM (\secref{prompt-design} below). The expression is validated through a
closed loop program synthesis step which checks that it is well-formed and evaluates to the target fragment
$s$. If this process succeeds, the system replaces the selected substring $s$ with the interpolation
expression \kw{\{$e$\}}, creating a new document configuration. This is the primary successful path through
the workflow and returns the system to the entry state where it is waiting for another top-level interaction
from the author.

Otherwise, the Fluid command-line either returned an error (either a syntax error or runtime error), or
succeeded but $e$ evaluated to some other string $s' \neq s$. Both cases trigger prompt augmentation with an
appropriate error message, and the system retries generation. Failure of the code synthesis step means that,
even after multiple retries with an augmented prompt, $e$ remained invalid or evaluated to $s' \neq s$. If $e$
was invalid (resulting in an error), no remedial action is open to the author. This is considered an
unsuccessful path through the workflow and returns the system to the start state. In the $s' \neq s$ case, the
can choose to manually abort and return to the entry state, or optionally to \emph{revise the goal}, replacing
$s$ with $s'$ in the target document and retaining $e$ as the candidate expression. This is intended to cover
the situation where the author has made a claim which is \emph{incorrect}, but the data set and surrounding
natural language has caused the LLM to synthesise an expression which generates a different value from the one
specified by the user.

\paragraph{Manual validation step.} Once a valid expression is approved, the updated text is spliced back into
the document, and any downstream interactions (e.g. reloading the web page) are checked.

This human-in-the-loop design combines automated synthesis with validation and author oversight, providing a
substantial level of automation, but requiring the author to intervene at key steps to ensure correctness.

@@
